The French domestic Pass Sanitaire did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals

### 2022-04-11 clean/code/vaccination-indicators.Rmd
clean up files

---
title: "Vaccination"
author: "FD"
output:
  html_document:
    self_contained: no
    keep_md: true
    toc: true
    toc_float: true
  pdf_document: default
---

@@

```{r loadVaccinationData}
# Source vaccination data
source("load-clean_vaccination.R")
```

@@

```{r}
save(totPopSize, totPopSize.adults, 
     nIndicatorsTot, nIndicatorsAfterRemovingCorrMiss, 
     datesAnalysis,
     nlocalities, 
     nrep,
     date.SP.announcement, date.SP.implementation, date.VP.announcement, date.VP.implementation, 
     dicPred, dic.fullpred,
     vacc,
     epci.PLM,
     txVaccAdults, 
     txVaccTot,
     popSizeDistrib, 
     pops, 
     maxOR.Unemployment_Benef,
     file = "savevalues.RData")

@@

# Essais

```{r, eval = FALSE}

v <- dat.all[, c("codgeo", "Abstention")]
vx <- v[which(is.na(v$Abstention)), ]
vx
v[1, ]
head(v)
tail(v)

library(inseeLocalData)
?inseeLocalData

#install.packages("insee")
library(insee)

get_dataset_list()

https://api.insee.fr/donnees-locales/V0.1/donnees/geo-POP@GEO2021RP2018/EPCI-247100647.INATC

```{r, eval = FALSE}
?prcomp

# Predictors
# New dataset with the predictors
dat2 <- dat.nocorr[, -1]

# For each predictor
for(col in colnames(dat2)){
  # Get the subset of the data
  v <- dat2[, col]
  # Compute the mean value, excluding NAs
  mv <- mean(v, na.rm = TRUE)
  # Fill in NAs with the mean
  v[is.na(v)] <- mv
  # Put back in the table
  dat2[, col] <- v
}

pca <- prcomp(dat2, center = TRUE, scale = TRUE)
ggbiplot(pca)

head(pca)

library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)

plot(pca$x[, c("PC1", "PC2")])

pcax <- as.data.frame(pca$x)
pcax$codgeo <- dat.nocorr$codgeo

thedate <- date3
sub <- vacc[which(vacc$date == thedate & vacc$classe_age != "TOUT_AGE"), ]
subb <- adultVacc(sub)

sub3 <- merge(subb, pcax, all = TRUE, by = "codgeo")
sub3$typeTaux <- discretizeQ(sub3$taux_cumu, prbs = seq(0, 1, by = 0.1))
palCat <- rev(met.brewer("Hiroshige", n = 10, type = "continuous"))
names(palCat) <- 1:10

plot(sub3$PC1, sub3$PC2, col = palCat[sub3$typeTaux], pch = 16)

str(pca)
head(sub3)
dim(subb)
dim(pcax)

sort(pca$rotation[, "PC1"], decreasing = TRUE)[1:30]
sort(pca$rotation[, "PC2"], decreasing = TRUE)[1:30]

head(pca)
summary(pca)

pca$sdev / sum(pca$sdev)
str(pca)

length(pca$rotation[, "PC2"])
length(subb$taux_cumu)

plot(sub3$PC2, sub3$taux_cumu)

out <- matrix(NA, nrow = (ncol(pcax) - 1), ncol = 4)
for(i in 1:(ncol(pcax) - 1)){
  mdl <- cor.test(sub3[, paste0("PC", i)], sub3$taux_cumu)
  out[i, ] <- c(mdl$estimate, mdl$conf.int, mdl$p.value)
}
out <- as.data.frame(out)
names(out) <- c("estimate", "ci1", "ci2", "pval")
out$PC <- seq_len(nrow(out))

out[order(abs(out$estimate), decreasing = TRUE), ][1:5, ]
ii <- out[order(abs(out$estimate), decreasing = TRUE), ][1:5, "PC"]

# Composition of the PCs
for(i in ii){
  print(i)
  print(t(sort(pca$rotation[, paste0("PC", i)], decreasing = TRUE)[1:20])) 
}
sort(pca$rotation[, "PC2"], decreasing = TRUE)[1:30]

plot(sort(abs(out$estimate)))

range(out$estimate)

head(out)
mdl2 <- cor.test(sub3$PC2, sub3$taux_cumu)
mdl2

mdl1 <- cor.test(sub3$PC1, sub3$taux_cumu)
mdl1
str(mdl1)
summary(mdl1)
mdl3 <- cor.test(sub3$PC3, sub3$taux_cumu)
mdl3

summary(mdl)

library(nlme)
?gls
         gls(divspe ~ loggrad_urb+grad_agrimean, method = "ML", 
corr=corExp(c(300000,0.7), form=~x_lambert93+y_lambert93, nugget=T), 
na.action = na.omit, data = df)

>>>>

---
title: "Vaccination"
author: "FD"
output: html_notebook:
  keep_md: true
editor_options: 
  chunk_output_type: console
---

@@

```{r loadVaccinationData}
# Source vaccination data
source("1_load-clean_vaccination.R")
```

@@

```{r}
save(totPopSize, totPopSize.adults, 
     nIndicatorsTot, nIndicatorsAfterRemovingCorrMiss, 
     datesAnalysis,
     nlocalities, 
     nrep,
     date.SP.announcement, date.SP.implementation, date.VP.announcement, date.VP.implementation, 
     dicPred, dic.fullpred,
     vacc,
     epci.PLM,
     txVaccAdults, 
     txVaccTot,
     popSizeDistrib, 
     pops, 
     maxOR.Unemployment_Benef,
     file = "outsavevalues.RData")

@@

### 2022-04-11 clean/ms/ms.Rmd
clean up files

load("../code/savevalues.RData")
load("../code/data_indicators.RData") 
```

@@

Attitudes toward vaccination are also known to be influenced by social and territorial inequalities. Surveys conducted in 2020 in France showed that respondents with lower education [@schwarzingerCOVID19VaccineHesitancy2021; @coulaudCOVID19VaccineIntention2022], lower income levels or less trust in authorities [@spireSocialInequalitiesHostility2021; @lindholtPublicAcceptanceCOVID192021] were more likely to be hostile to COVID-19 vaccines. <!-- A survey conducted in July 2021 in France [@bajosWhenMistrustGovernment2022] on close to 81000 participants indicated that unvaccinated respondents were more likely to have lower income and more likely to belong to racialised minorities than vaccinated ones. The survey also confirmed that mistrust in the government was fueling refusal to be vaccinated [@bajosWhenMistrustGovernment2022]. -->

<!-- XX MOVE INTERNATIONAL COMPARISONS IN THE DISCUSSION Whether differential intentions translated into effective differences in vaccination, as described in other countries [Caspi21, Murthy21], remained to be shown in France.  -->

@@

The vaccination dataset for mainland France encompasses about `r round(pops[nrow(pops), 1] / 10^6, 1)` million individuals (median district size `r format(round(popSizeDistrib["50%"]), scientific = FALSE)` inhabitants, interquartile range `r format(round(popSizeDistrib["25%"]), scientific = FALSE)`--`r format(round(popSizeDistrib["75%"]), scientific = FALSE)`). The vaccination data are available by age class: 00--19, 20--39, 40--54, 55--64, 65--74, 75 and over. Population sizes for each locality and each age class are also provided. <!-- These population sizes are updated monthly to account for individuals aging, potential residential moves, and deaths. XXX TODO: Demander pourquoi les tailles de pop sont constantes au cours du temps FD: Removed because was not specified -->

@@

<!--

@@

<!-- [moved to Discussion]
The share of local income coming from unemployment benefits (`Unemployment_Benef`) and the proportion of overcrowded households (`Overcrowding_rate`) are two indicators linked to poverty. Votes for Mr Asselineau (`Asselineau`) represented a minority of cast votes in 2017 in France (less than $1\%$ overall). This candidate for instance proposed that France exits the European Union, leave the Euro zone and reinstall the Franc currency; he was a proponent of hydroxychloroquin and ivermectin during the Covid-19 pandemic, and publicly expressed doubts about the safety of available Covid-19 vaccines. 
-->

@@

```{r figMap, fig.height=5, fig.width=4, fig.cap="Adult vaccination rates by district of mainland France", echo = FALSE}
source("../code/2_plot-map.R")
source("../code/1_functions-adjustAges.R")
load("../data/mapFiles_withDepReg.RData")
load("../data/chefslieux.RData")
# Choose date
thedate <- datesAnalysis[6]
# Vaccination data by age at that date
tmpVacc <- vacc[vacc$date == thedate & vacc$classe_age != "TOUT_AGE", ]
#tmpVacc2 <- vacc[vacc$date == thedate & vacc$classe_age == "TOUT_AGE", ]
#tmpVacc2$taux_cumu <- tmpVacc2$taux_cumu_1_inj

@@

Our results, based on exhaustive national datasets, indicate that the French sanitary pass, and the later vaccine pass, did not solve Covid-19 vaccination heterogeneities, but instead crystallized them. <!-- Indicators that were the most strongly associated to vaccination rates are linked to poverty (the share of local income coming from unemployment benefits (`Unemployment_Benef`), and the proportion of overcrowded households (`Overcrowding_rate`)), immigration (the proportion of immigrants in the population (`Immigrant`)), -->
Indicators most associated with vaccination rates were associated to poverty, immigration, anti-establishment vote (or abstention), and a North-West -- South-East contrast. For instance, the odds for an adult to still be unvaccinated by the end of February 2021 are about `r round(maxOR.Unemployment_Benef, 1)` times higher when living in the districts with higher than median value share of income coming from unemployment benefits, than when living in the districts with lower than median value.

>>>>

load("../code/outsavevalues.RData")
load("../code/data_indicators.RData") 
```

@@

Attitudes toward vaccination are also known to be influenced by social and territorial inequalities. Surveys conducted in 2020 in France showed that respondents with lower education [@schwarzingerCOVID19VaccineHesitancy2021; @coulaudCOVID19VaccineIntention2022], lower income levels or less trust in authorities [@spireSocialInequalitiesHostility2021; @lindholtPublicAcceptanceCOVID192021] were more likely to be hostile to COVID-19 vaccines.

@@

The vaccination dataset for mainland France encompasses about `r round(pops[nrow(pops), 1] / 10^6, 1)` million individuals (median district size `r format(round(popSizeDistrib["50%"]), scientific = FALSE)` inhabitants, interquartile range `r format(round(popSizeDistrib["25%"]), scientific = FALSE)`--`r format(round(popSizeDistrib["75%"]), scientific = FALSE)`). The vaccination data are available by age class: 00--19, 20--39, 40--54, 55--64, 65--74, 75 and over. Population sizes for each locality and each age class are also provided.

@@

<!-- 
Key dates:

@@

@@

```{r figMap, fig.height=5, fig.width=4, fig.cap="Adult vaccination rates by district of mainland France", echo = FALSE}
source("../code/2_plot-map.R")
source("../code/1_functions-adjustAges.R")
load("../data/mapFiles_withDepReg.RData")
load("../data/mapchefslieux.RData")
# Choose date
thedate <- datesAnalysis[6]
# Vaccination data by age at that date
tmpVacc <- vacc[vacc$date == thedate & vacc$classe_age != "TOUT_AGE", ]
#tmpVacc2 <- vacc[vacc$date == thedate & vacc$classe_age == "TOUT_AGE", ]
#tmpVacc2$taux_cumu <- tmpVacc2$taux_cumu_1_inj

@@

Our results, based on exhaustive national datasets, indicate that the French sanitary pass, and the later vaccine pass, did not solve Covid-19 vaccination heterogeneities, but instead crystallized them. 
Indicators most associated with vaccination rates were associated to poverty, immigration, anti-establishment vote (or abstention), and a North-West -- South-East contrast. For instance, the odds for an adult to still be unvaccinated by the end of February 2021 are about `r round(maxOR.Unemployment_Benef, 1)` times higher when living in the districts with higher than median value share of income coming from unemployment benefits, than when living in the districts with lower than median value.
